Carbonate chemistry at Ny-Ålesund

### 2021-12-22 awipev-CO2 (original).Rmd
dunno

---
title: "Carbonate chemistry at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane and Bernard Gentili"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  word_document: default
  pdf_document: default
  odt_document: default
editor_options: 
  chunk_output_type: console
---
![ ](Images/AWIPEV-Logo.jpg){width=250px}

@@

#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) ,
                hour = lubridate::hour(d_hour$datetime)# abbreviated months
  )
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all)))
rm(d_hour)
```

@@

```{r read and prepare discrete AT+CT, echo=FALSE, warning=FALSE, message=FALSE}
d <- read_excel(paste0(path, "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.xlsx"),
                col_names = TRUE) %>% 
#   d <- tibble(
#   read.table(paste0(path, "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.csv"), header = T, dec = ".", as.is = T, sep = ";", fill = TRUE)
# ) %>% 
  dplyr::mutate(datetime = dmy_hm(sampling_date, tz="UTC"),
                measure_date = dmy(measure_date, tz="UTC"),
                date = as.Date(datetime)) %>% 
  #dplyr::filter(datetime < as.POSIXct("2021-01-01 00:00:00")) %>% #eliminate 2021 data
  dplyr::rename(ph_s_sf = pH_s_seafet,
                ph_s_dur = pH_s_durafet,
                qflag_ph_s_sf = qflag_pH_s_seafet,
                qflag_ph_s_dur = qflag_pH_s_durafet)

# Add the flag location: some samples have been collected at the pier. Notes them here.
# Flag 1 = collected on the pIer by Niskin and flag 0 = collected in the FB (normal collect)
 # And keep data with flag 0 for FB location
d <- d %>%
  dplyr::mutate(location_flag = ifelse(datetime == as.POSIXct("2016-04-21 09:25:00") | 
                                         datetime == as.POSIXct("2016-02-10 14:00:00") | 
                                         datetime == as.POSIXct("2016-02-17 09:20:00") |
                                         datetime == as.POSIXct("2016-02-24 14:20:00") |
                                         datetime == as.POSIXct("2016-02-04 09:30:00") |
                                         datetime == as.POSIXct("2016-01-27 15:00:00") |
                                         datetime == as.POSIXct("2016-01-13 13:45:00") |
                                         datetime == as.POSIXct("2016-01-06 14:10:00") |
                                         datetime == as.POSIXct("2015-12-24 10:45:00") |
                                         datetime == as.POSIXct("2015-12-30 13:30:00") |
                                         datetime == as.POSIXct("2018-01-05 15:20:00") |
                                         datetime == as.POSIXct("2018-02-02 14:35:00") |
                                         datetime == as.POSIXct("2018-03-02 14:10:00") |
                                         datetime == as.POSIXct("2019-05-31 11:45:00"), 
                                       1, 0))
d <- d %>%
  dplyr::filter(location_flag == 0) %>% 
# below we change the time of 2 discrete samplings during period 1 of seafet deployment
# This is to enable calibration when the REMOS was at 9 m depth
  dplyr::mutate(datetime = case_when(datetime == as.POSIXct("2017-08-24 13:30:00") ~ as.POSIXct("2017-08-24 19:00:00"),
                                     datetime == as.POSIXct("2017-11-28 09:15:00") ~ as.POSIXct("2017-11-28 13:00:00"),
                   TRUE ~ datetime))

@@

```{r read and prepare nutrient data, echo=FALSE, warning=TRUE, message=FALSE}
# File "Ferrybox nutrients 2014-2018.xlsx" provided by 
# Willem van de Poll <w.h.van.de.poll@rug.nl> on 2020-01-16
# nutrients in umol/l
nuts <- readxl::read_excel(paste0(path, "fb_data/Ferrybox nutrients 2014-2018.xlsx"),
                           skip =1) %>% 
  drop_na(date) %>% 
  dplyr::mutate(datetime = as.POSIXct(paste0(date, " 12:00:00")),
                ) %>% # assuming sampling time 12:00
  dplyr::filter(datetime > as.POSIXct("2015-07-09"))
# The nearest times
time_index <- unique(d_all$datetime)
# nuts$closest <- time_index[
#   FNN::knnx.index(time_index, # The values to be queried (big df)
#                   as.vector(nuts$datetime), # The value to find the closest match to (small df)
#                   k = 1)] # The number of closest values to return
nuts <- nuts %>% 
  dplyr::mutate(closest = time_index[
  FNN::knnx.index(time_index, # The values to be queried (big df)
                  as.vector(nuts$datetime), # The value to find the closest match to (small df)
                  k = 1)]) %>%  # The number of closest values to return
  dplyr::select(-date, -datetime) %>% 
  dplyr::rename(NO3NO2 = `NO3+NO2`)
# cleanup and convert in umol/kg
d_all <- left_join(d_all, nuts, by = c("datetime" = "closest")) #%>% 
  #dplyr::select(-closest_datetime)
d_all <- d_all %>% 
  dplyr::mutate(density_fb = seacarb::rho(S = sal_fb, T = temp_fb, P = 0),
                po4 = PO4 / (density_fb/1000),
                nh4 = NH4 / (density_fb/1000),
                no3no2 = NO3NO2 / (density_fb/1000),
                no3 = NO3 / (density_fb/1000),
                si = Si / (density_fb/1000)
  ) %>% 
  dplyr::select(-PO4, -NH4, -NO3NO2, -NO3, -Si)
```

@@

```{r plot measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
print(p$plot)
rm(fit, p)
```

@@

Only 5 spectrophotometric pH which could be used for calibration when seaFET at > 8 m: 2018-11-16, 2019-04-17, 2019-10-11, 2020-01-27. All deployment meriods are covered but we had to alter the sampling time of:

* 2017-08-24 14:00 to 2017-08-24 19:00 (5 h later)
* 2017-11-28 09:00 to 2017-11-28 13:00 (4 h later)

EO is calculated making a mean of several E0 that come from one deployment period (sensor in the water for several months).

@@

```{r seaFET calibration preparation, echo=FALSE, message= FALSE, warning=FALSE, out.width= "60%"}
sfinterv <- as.POSIXct(c("2017-08-24 12:00:00","2018-04-17 12:00:00", "2019-09-02 12:00:00"))

# add here the new periods sfinterv[..]
d_all <- d_all %>%
  dplyr::mutate(deployment_sf = ifelse(datetime >= sfinterv[1] & datetime <= sfinterv[2], 01,
                                ifelse(datetime >= sfinterv[2] & datetime <= sfinterv[3], 02,
                                ifelse(datetime >= sfinterv[3], 03, NA ))))
# dygraph to check seafet data
# pH_xts <-
#   dplyr::select(d_all, datetime, voltINT, voltEXT, sal_fb, temp_insitu_ctd, ph_s_sf_temp_insi)
# pH_xts <-
#   as.xts(pH_xts, order.by = d_all$datetime)
# dygraph(pH_xts,
#         group = "awipev",
#         main = "pH",
#         ylab = "pH") %>%
#   dySeries(
#     axis = "y2",
#     "voltINT",
#     label = "voltINT",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[1],
#     strokeWidth = 0,
#     pointSize = 2
#   ) %>%
#   dySeries(axis="y2",
#     "voltEXT",
#     label = "voltEXT",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[2],
#     strokeWidth = 0,
#     pointSize = 2
#   ) %>%
#    dySeries(
#     "temp_insitu_ctd",
#     label = "temp_insitu_ctd",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[3],
#     strokeWidth = 0,
#     pointSize = 2
#   ) %>%
#    dySeries(
#     "sal_fb",
#     label = "sal_fb",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[4],
#     strokeWidth = 0,
#     pointSize = 2
#   ) %>%
#    dySeries(
#     "ph_s_sf_temp_insi",
#     label = "ph_s_sf_temp_insi",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[5],
#     strokeWidth = 0,
#     pointSize = 4
#   ) %>%
#   dyHighlight(
#     highlightCircleSize = 8,
#     highlightSeriesBackgroundAlpha = 0.5,
#     hideOnMouseOut = TRUE
#   ) %>%
#   dyLegend(show = "follow")   %>%
#   dyAxis("y2",valueRange = c(-2,2))%>%
#   dyOptions(
#     drawGrid = TRUE,
#     drawPoints = TRUE,
#     pointSize = c(1, 8, 1),
#     useDataTimezone = TRUE
#   ) %>%
#   dyRangeSelector(height = 30, dateWindow = NULL)

# To make sf calibration against pH spectro, we need: sal - T insi - voltINT/EXT . If we missed one of these parameters, calibration will not run
# Here we try to fill the gap when it is possible: make a mean with 4 previous/next hours available in the time serie. 
# If the pH spectro is too far from a needed value, we decided to skip it for the calibration.

# (1) Raw voltage INT + pHINT must be filtered from 2017-12-14 to 2018-01-08 < 8.04,
# d_all$phINT <- ifelse(d_all$datetime >= as.POSIXct("2017-12-14") &
#                         d_all$datetime <= as.POSIXct("2018-01-08") &
#                         d_all$phINT < 8.04, NA, d_all$phINT)
# d_all$phEXT <-ifelse(d_all$datetime >= as.POSIXct("2017-12-14") &
#                        d_all$datetime <= as.POSIXct("2018-01-08") &
#                        d_all$phEXT < 8.04, NA, d_all$phEXT) 
d_all$voltINT <-ifelse(d_all$datetime >= as.POSIXct("2017-12-14") &
                         d_all$datetime <= as.POSIXct("2018-01-08") &
                         d_all$voltINT < -0.9439, NA, d_all$voltINT) 
d_all$voltEXT <-ifelse(d_all$datetime >= as.POSIXct("2017-12-14") &
                         d_all$datetime <= as.POSIXct("2018-01-08") &
                         d_all$voltEXT < -0.907, NA, d_all$voltEXT) 
#(2) removed from 2018-01-08 to 2018-02-02 due to drift
d_all$voltINT <- ifelse(d_all$datetime >= as.POSIXct("2018-01-08 17:00:00") &
                          d_all$datetime <= as.POSIXct("2018-02-02 09:00:00"), 
                        NA, d_all$voltINT)
d_all$voltEXT <- ifelse(d_all$datetime >= as.POSIXct("2018-01-08 17:00:00") &
                          d_all$datetime <= as.POSIXct("2018-02-02 09:00:00"),
                        NA, d_all$voltEXT)
d_all$phINT <- ifelse(d_all$datetime >= as.POSIXct("2018-01-08 17:00:00") &
                        d_all$datetime <= as.POSIXct("2018-02-02 09:00:00"),
                      NA, d_all$phINT)
d_all$phEXT <- ifelse(d_all$datetime >= as.POSIXct("2018-01-08 17:00:00") &
                        d_all$datetime <= as.POSIXct("2018-02-02 09:00:00"),
                      NA, d_all$phEXT)
# (3) outliers removed
d_all$voltINT <- ifelse(d_all$datetime == as.POSIXct("2017-11-02 21:00:00") |
                          d_all$datetime == as.POSIXct("2018-04-18 14:00:00"),
                        NA, d_all$voltINT)

@@

### add the depth into discrete df. In order to see where the seafet was.
#discrete <- left_join(discrete, d_all%>%dplyr::select(closest_datetime,depth, depth_discrete),by="closest_datetime")
discrete <- left_join(discrete, d_all%>%dplyr::select(closest_datetime,depth),by="closest_datetime")
discrete <- discrete%>%dplyr::filter(discrete$datetime < "2021-07-01 00:00:00")

#offset_xts <- select(discrete, closest_datetime, depth,depth_discrete, ph_s_sf_temp_insi)
offset_xts <- select(discrete, closest_datetime, depth, ph_s_sf_temp_insi)
offset_xts <- as.xts(offset_xts, order.by=offset_xts$closest_datetime)

dygraph(offset_xts, main = "") %>%
#dySeries("depth_discrete", drawPoints = TRUE, pointSize=3, strokeWidth = 1,label = "depth_discrete", color="black")%>%
dySeries("depth", drawPoints = TRUE, pointSize=3, label = "depth", strokeWidth = 1,color="blue")%>%
dySeries(axis="y2","ph_s_sf_temp_insi", drawPoints = TRUE, pointSize=4, label = "spectro", color="red")%>%
    dyLegend(show = "follow")   %>%
    dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
dyOptions(
    drawGrid = TRUE,
    strokeWidth= 0,
    drawPoints = TRUE,
    useDataTimezone = TRUE,
    digitsAfterDecimal = 4) %>%
  dyRangeSelector(height = 30, dateWindow= c("2017-10-21 00:00:00", "2021-01-11 00:00:00"))
```

```{r seaFET pH sf_calib , echo=FALSE, message= FALSE, warning=FALSE}
## calculation of E0INT25/E0EXT25 irrespective of the depth at which the seafet is (all data used)
calib_spectro_all <- sf_calib(calEint = d_all$voltINT, calEext = d_all$voltEXT, calpH = d_all$ph_s_sf_temp_insi, calT = d_all$temp_sf, calSal = d_all$sal_insitu_ctd_interp)
# put number of deployment 
calib_spectro_all$deployment_sf <- d_all$deployment_sf

# make the mean of E0 by period of deployement
calib_spectro_all <- group_by(calib_spectro_all, deployment_sf) %>%
  dplyr::summarise(E0int25 = mean(E0int25, na.rm = TRUE),
            E0ext25 = mean(E0ext25, na.rm = TRUE) 
            ) %>% 
  dplyr::filter(!is.na(deployment_sf))
names(calib_spectro_all) <- c("deployment_sf", "E0int25_sf", "E0ext25_sf")

# add E0 mean to seaFET data per deployment
d_all <-left_join(d_all, calib_spectro_all, by="deployment_sf" )

## calculation of E0INT25/E0EXT25 only using seafet data deeper than 8m
d_all_calib_10m <- d_all %>%
#  dplyr::select(datetime, voltINT, voltEXT,ph_s_sf_temp_insi,sal_insitu_ctd_interp,temp_sf, depth_discrete,deployment_sf)%>%
  dplyr::select(datetime, voltINT, voltEXT,ph_s_sf_temp_insi,sal_insitu_ctd_interp,temp_sf,depth, deployment_sf)%>%
#  dplyr::filter(depth_discrete == 10)
    dplyr::filter(depth > 8)

calib_spectro_10m <- sf_calib(calEint = d_all_calib_10m$voltINT, calEext = d_all_calib_10m$voltEXT, calpH = d_all_calib_10m$ph_s_sf_temp_insi, calT = d_all_calib_10m$temp_sf, calSal = d_all_calib_10m$sal_insitu_ctd_interp)
# put number of deployment 
calib_spectro_10m$deployment_sf <- d_all_calib_10m$deployment_sf

# make the mean of E0 by period of deployement
calib_spectro_10m <- group_by(calib_spectro_10m, deployment_sf) %>%
  summarise(E0int25 = mean(E0int25, na.rm = TRUE),
            E0ext25 = mean(E0ext25, na.rm = TRUE) 
            ) %>% 
  dplyr::filter(!is.na(deployment_sf))
names(calib_spectro_10m) <- c("deployment_sf", "E0int25_sf_10m", "E0ext25_sf_10m")

# add E0 mean to seaFET data per deployment
d_all <- dplyr::left_join(d_all, calib_spectro_10m, by="deployment_sf")
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all)))
```

Below are the seafet calibration parameters using all data (irrespective of the depth of the seafet) and with the seafet data collected below 8 m. They are very close but we nevertheless use the calibration data below 8 m (easier to explain in the manuscript).

```{r seaFET pH plot sf_calib , echo=FALSE, message= FALSE, warning=FALSE}
#### plot
# kk <- tidyr::pivot_longer(d_all, cols = starts_with("E0"), names_to = "variable", values_to = "value")
# p_calib <- ggplot(kk) + geom_point(aes(x=deployment_sf, y=value, color=variable), size=2) +
#   facet_grid(variable ~ ., scales="fixed") + theme_bw() +
#   scale_colour_discrete(guide="none")

kk <- d_all %>% 
  dplyr::select(deployment_sf, tidyr::starts_with("E0")) %>% 
  tidyr::pivot_longer(cols = starts_with("E0"), names_to = "variable", values_to = "value")
p_calib <- ggplot(kk) + geom_point(aes(x=deployment_sf, y=value, color=variable), size=2) +
  facet_grid(variable ~ ., scales="fixed") + theme_bw() +
  scale_colour_discrete(guide="none")
print(p_calib)

calib_spectro <- cbind(calib_spectro_all, calib_spectro_10m)
kbl(calib_spectro)%>%
    kable_paper() %>% 
    kable_styling(fixed_thead = T)
```

### **seaFET pHT calibrated**
Final calibrated seaFET pHT data (small green dots) are shown here after calibration with spectro measurements (red dots) according to each period of deployement, defined between 2 spectro measurements (shading grey).

@@

#- pH INT - pH EXT was plotted here.

>>>>

---
title: "Carbonate chemistry at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  word_document: default
  pdf_document: default
  odt_document: default
editor_options: 
  chunk_output_type: console
---
![ ](Images/AWIPEV-Logo.jpg){width=250px}

@@

#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) ,
                hour = lubridate::hour(d_hour$datetime)# abbreviated months
  )
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all)))
rm(d_hour)

@@

```{r read and prepare discrete AT+CT, echo=FALSE, warning=FALSE, message=FALSE}
d <- read_excel(paste0(path, "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.xlsx"),
                col_names = TRUE) %>% 
#   d <- tibble(
#   read.table(paste0(path, "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.csv"), header = T, dec = ".", as.is = T, sep = ";", fill = TRUE)
# ) %>% 
  dplyr::mutate(datetime = dmy_hm(sampling_date, tz="UTC"),
                measure_date = ymd(measure_date, tz="UTC"),
                date = as.Date(datetime)) %>% 
  #dplyr::filter(datetime < as.POSIXct("2021-01-01 00:00:00")) %>% #eliminate 2021 data
  dplyr::rename(ph_s_sf = pH_s_seafet,
                ph_s_dur = pH_s_durafet,
                qflag_ph_s_sf = qflag_pH_s_seafet,
                qflag_ph_s_dur = qflag_pH_s_durafet)

# Add the flag location: some samples have been collected at the pier. Notes them here.
# Flag 1 = collected on the pIer by Niskin and flag 0 = collected in the FB (normal collect)
 # And keep data with flag 0 for FB location
d <- d %>%
  dplyr::mutate(location_flag = ifelse(datetime == as.POSIXct("2016-04-21 09:25:00") | 
                                         datetime == as.POSIXct("2016-02-10 14:00:00") | 
                                         datetime == as.POSIXct("2016-02-17 09:20:00") |
                                         datetime == as.POSIXct("2016-02-24 14:20:00") |
                                         datetime == as.POSIXct("2016-02-04 09:30:00") |
                                         datetime == as.POSIXct("2016-01-27 15:00:00") |
                                         datetime == as.POSIXct("2016-01-13 13:45:00") |
                                         datetime == as.POSIXct("2016-01-06 14:10:00") |
                                         datetime == as.POSIXct("2015-12-24 10:45:00") |
                                         datetime == as.POSIXct("2015-12-30 13:30:00") |
                                         datetime == as.POSIXct("2018-01-05 15:20:00") |
                                         datetime == as.POSIXct("2018-02-02 14:35:00") |
                                         datetime == as.POSIXct("2018-03-02 14:10:00") |
                                         datetime == as.POSIXct("2019-05-31 11:45:00"), 
                                       1, 0))
d <- d %>%
  dplyr::filter(location_flag == 0)

@@

```{r read and prepare nutrient data, echo=FALSE, warning=TRUE, message=FALSE}
# File "Ferrybox nutrients 2014-2018.xlsx" provided by 
# Willem van de Poll <w.h.van.de.poll@rug.nl> on 2020-01-16
# nutrients in umol/l
nuts <- readxl::read_excel(paste0(path, "fb_data/Ferrybox nutrients 2014-2018.xlsx"),
                           skip =1) %>% 
  drop_na(date) %>% 
  dplyr::mutate(datetime = as.POSIXct(paste0(date, " 12:00:00")),
                ) %>% # assuming sampling time 12:00
  dplyr::filter(datetime > as.POSIXct("2015-07-09"))
# The nearest times
time_index <- unique(d_all$datetime)
# nuts$closest <- time_index[
#   FNN::knnx.index(time_index, # The values to be queried (big df)
#                   as.vector(nuts$datetime), # The value to find the closest match to (small df)
#                   k = 1)] # The number of closest values to return
nuts <- nuts %>% 
  dplyr::mutate(closest = time_index[
  FNN::knnx.index(time_index, # The values to be queried (big df)
                  as.vector(nuts$datetime), # The value to find the closest match to (small df)
                  k = 1)]) %>%  # The number of closest values to return
  dplyr::select(-date, -datetime) %>% 
  dplyr::rename(NO3NO2 = `NO3+NO2`)
# cleanup and convert in umol/kg
d_all <- left_join(d_all, nuts, by = c("datetime" = "closest")) #%>% 
  #dplyr::select(-closest_datetime)
d_all <- d_all %>% 
  dplyr::mutate(density_fb = seacarb::rho(S = sal_fb, T = temp_fb, P = 0),
                po4 = PO4 / (density_fb/1000),
                nh4 = NH4 / (density_fb/1000),
                no3no2 = NO3NO2 / (density_fb/1000),
                no3 = NO3 / (density_fb/1000),
                si = Si / (density_fb/1000)
  )
```

@@

```{r plot measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
print(p$plot)
# 
```

@@

Spectrophotometric pH which could be used for calibration: 2017-08-24, 2017-09-28, 2019-09-13 x 2.
Spectrophotometric pH which could not be used for calibration : 2018-12-14, 2019-01-18.

We decided to change the calculation of E0. Before E0 was calculated for each pH spectro measurements. Now, EO is calculated making a mean of several E0 that come from one deployment period (sensor in the water for several months).

@@

```{r seaFET calibration sf_calib, echo=FALSE, message= FALSE, warning=FALSE, out.width= "60%"}

# # We decided to calibrate pH data within each pH spectro measurement.
# # Thus, each pH spectro is taken as one calibration periode. 
# # We calculate here the day between 2 pH spectro in order to have the start/end of one deployment period.
# sfPeriod <- d_all%>%
#   dplyr::select(datetime, ph_s_sf_temp_insi)%>%
#   dplyr::filter(!is.na(ph_s_sf_temp_insi))
# # Remove pH spectro that not allow a calibration because of missing calibration value (voltINT, Sal or Temp insi)
# sfPeriod <- sfPeriod%>%
# dplyr::filter(datetime != "2018-04-16 07:00:00" & datetime != "2018-12-14 09:00:00" & datetime != "2019-01-18 09:00:00")
# 
# sfPeriod <-  sfPeriod %>%
#   dplyr::arrange(datetime) %>%
#   dplyr::mutate(middle=  seconds_to_period(c(0,difftime(time1= datetime[-1], time2= datetime[-nrow(sfPeriod)])/2* 3600)))%>%
#   dplyr::mutate(sf_interval = datetime + middle[2:nrow(sfPeriod)])
# 
# #put interval in a list with the start date of seaFET measurements
# sfinterv <- c( as_datetime("2017-08-24 13:00:00"), sfPeriod[,4])
# #Change period sfinterv[5]. make the end of this period longer until 2018-03-21 23:59:00 
# # This is needed in order to take in consideration pH from March 2018 with a calibration E0 lower (-1.4) than the next period (-0.38).
# sfinterv[5] <- ymd_hms("2018-03-22 23:59:59")
# #  #Change period sfinterv[6]. Because seafet was off and on on the 2018-04-24 19:00:00
# # sfinterv[6] <- ymd_hms("2018-04-24 18:00:00")
# #  #Change period sfinterv[6]. Because seafet was off and on on the 2018-04-24 19:00:00
# # sfinterv[14] <- ymd_hms("2019-01-29 16:00:00")
# 
# #sfinterv <-sort(unlist(sfinterv, use.names=FALSE))
# # set deployment period between 2 spectro measurements.

# # add here the new periods sfinterv[..]
# d_all <- d_all %>%
#   dplyr::mutate(deployment_sf = ifelse(datetime >= sfinterv[1] & datetime <= sfinterv[2], 01, 
#                                 ifelse(datetime >= sfinterv[2] & datetime <= sfinterv[3], 02, 
#                                 ifelse(datetime >= sfinterv[3] & datetime <= sfinterv[4], 03, 
#                                 ifelse(datetime >= sfinterv[4] & datetime <= sfinterv[5], 04,
#                                 ifelse(datetime >= sfinterv[5] & datetime <= sfinterv[6], 05,
#                                 ifelse(datetime >= sfinterv[6] & datetime <= sfinterv[7], 06,
#                                 ifelse(datetime >= sfinterv[7] & datetime <= sfinterv[8], 07,
#                                 ifelse(datetime >= sfinterv[8] & datetime <= sfinterv[9], 08,
#                                 ifelse(datetime >= sfinterv[9] & datetime <= sfinterv[10], 09,
#                                 ifelse(datetime >= sfinterv[10] & datetime <= sfinterv[11], 10,
#                                 ifelse(datetime >= sfinterv[11] & datetime <= sfinterv[12], 11,
#                                  ifelse(datetime >= sfinterv[12] & datetime <= sfinterv[13], 12,
#                                  ifelse(datetime >= sfinterv[13] & datetime <= sfinterv[14], 13,
#                                  ifelse(datetime >= sfinterv[14] & datetime <= sfinterv[15], 14,
#                                  ifelse(datetime >= sfinterv[15] & datetime <= sfinterv[16], 15,
#                                  ifelse(datetime >= sfinterv[16] & datetime <= sfinterv[17], 16,
#                                  ifelse(datetime >= sfinterv[17] & datetime <= sfinterv[18], 17,
#                                 ifelse(datetime >= sfinterv[18] & datetime <= sfinterv[19], 18,
#                                 ifelse(datetime >= sfinterv[19] & datetime <= sfinterv[20], 19,
#                                 ifelse(datetime >= sfinterv[20] & datetime <= sfinterv[21], 20,
#                                 ifelse(datetime >= sfinterv[21] & datetime <= sfinterv[22], 21,
#                                 #ifelse(datetime >= sfinterv[22] & datetime <= sfinterv[23], 22,
#                                 #ifelse(datetime >= sfinterv[23] & datetime <= sfinterv[24], 23,
#                                 #ifelse(datetime >= sfinterv[24] & datetime <= sfinterv[25], 24,
#                                 #ifelse(datetime >= sfinterv[25] & datetime <= sfinterv[26], 25,
#                                 ifelse(datetime >= sfinterv[22] , 22,NA )))))))))))))))))))))))

# set deployment period when seaFET is in/out of water
sfinterv <- as.POSIXct(c("2017-08-24 12:00:00","2018-04-17 12:00:00", "2019-09-02 12:00:00"))

# add here the new periods sfinterv[..]
d_all <- d_all %>%
  dplyr::mutate(deployment_sf = ifelse(datetime >= sfinterv[1] & datetime <= sfinterv[2], 01,
                                ifelse(datetime >= sfinterv[2] & datetime <= sfinterv[3], 02,
                                ifelse(datetime >= sfinterv[3], 03, NA ))))

# pco2_contros_xts <-
#   dplyr::select(d_all, datetime, voltINT, voltEXT, sal_fb, temp_insitu_ctd, ph_s_sf_temp_insi)
# pco2_contros_xts <-
#   as.xts(pco2_contros_xts, order.by = d_all$datetime)
# dygraph(pco2_contros_xts,
#         group = "awipev",
#         main = "<i>p</i>CO<sub>2</sub>",
#         ylab = "<i>p</i>CO<sub>2</sub>") %>%
#   dySeries(
#     axis = "y2",
#     "voltINT",
#     label = "voltINT",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[1],
#     strokeWidth = 0,
#     pointSize = 2
#   ) %>%
#   dySeries(axis="y2",
#     "voltEXT",
#     label = "voltEXT",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[2],
#     strokeWidth = 0,
#     pointSize = 2
#   ) %>%
#    dySeries(
#     "temp_insitu_ctd",
#     label = "temp_insitu_ctd",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[3],
#     strokeWidth = 0,
#     pointSize = 2
#   ) %>%
#    dySeries(
#     "sal_fb",
#     label = "sal_fb",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[4],
#     strokeWidth = 0,
#     pointSize = 2
#   ) %>%
#    dySeries(
#     "ph_s_sf_temp_insi",
#     label = "ph_s_sf_temp_insi",
#     color =  RColorBrewer::brewer.pal(5, "Dark2")[5],
#     strokeWidth = 0,
#     pointSize = 4
#   ) %>%
#   dyHighlight(
#     highlightCircleSize = 8,
#     highlightSeriesBackgroundAlpha = 0.5,
#     hideOnMouseOut = TRUE
#   ) %>%
#   dyLegend(show = "follow")   %>%
#   dyAxis("y2",valueRange = c(-2,2))%>%
#   dyOptions(
#     drawGrid = TRUE,
#     drawPoints = TRUE,
#     pointSize = c(1, 8, 1),
#     useDataTimezone = TRUE
#   ) %>%
#   dyRangeSelector(height = 30, dateWindow = NULL)

# To make sf calibration against pH spectro, we need: sal - T insi - voltINT/EXT . If we missed one of these parameters, calibration will not run
# Here we try to fill the gap when it is possible: make a mean with 4 previous/next hours available in the time serie. 
# If the pH spectro is too far from a needed value, we decided to skip it for the calibration.
# we highlighted 7 pH spectro that dont match one of these parameters : 2017-08-24 14:00:00 - 2017-09-28 13:00:00 - 2018-04-16 07:00:00 - 2018-12-14 09:00:00 - 2019-01-18 09:00:00 - 2019-09-13 08:00:00 - 2019-09-13 13:00:00
# (1) Raw voltage INT + pHINT must be filtered from 2017-12-14 to 2018-01-08 < 8.04,
# (2) removed from 2018-01-08 to 2018-02-02
# (3) outliers removed
# (1)
d_all$phINT <- ifelse(d_all$datetime >= as.POSIXct("2017-12-14") &
                        d_all$datetime <= as.POSIXct("2018-01-08") &
                        d_all$phINT < 8.04, NA, d_all$phINT)
d_all$phEXT <-ifelse(d_all$datetime >= as.POSIXct("2017-12-14") &
                       d_all$datetime <= as.POSIXct("2018-01-08") &
                       d_all$phEXT < 8.04, NA, d_all$phEXT) 
d_all$voltINT <-ifelse(d_all$datetime >= as.POSIXct("2017-12-14") &
                         d_all$datetime <= as.POSIXct("2018-01-08") &
                         d_all$voltINT < -0.9439, NA, d_all$voltINT) 
d_all$voltEXT <-ifelse(d_all$datetime >= as.POSIXct("2017-12-14") &
                         d_all$datetime <= as.POSIXct("2018-01-08") &
                         d_all$voltEXT < -0.907, NA, d_all$voltEXT) 
# (2)
d_all$voltINT <- ifelse(d_all$datetime >= as.POSIXct("2018-01-08 17:00:00") &
                          d_all$datetime <= as.POSIXct("2018-02-02 09:00:00"), 
                        NA, d_all$voltINT)
d_all$voltEXT <- ifelse(d_all$datetime >= as.POSIXct("2018-01-08 17:00:00") &
                          d_all$datetime <= as.POSIXct("2018-02-02 09:00:00"),
                        NA, d_all$voltEXT)
d_all$phINT <- ifelse(d_all$datetime >= as.POSIXct("2018-01-08 17:00:00") &
                        d_all$datetime <= as.POSIXct("2018-02-02 09:00:00"),
                      NA, d_all$phINT)
d_all$phEXT <- ifelse(d_all$datetime >= as.POSIXct("2018-01-08 17:00:00") &
                        d_all$datetime <= as.POSIXct("2018-02-02 09:00:00"),
                      NA, d_all$phEXT)
# (3)
d_all$voltINT <- ifelse(d_all$datetime == as.POSIXct("2017-11-02 21:00:00") |
                          d_all$datetime == as.POSIXct("2018-04-18 14:00:00"),
                        NA, d_all$voltINT)

@@

####
## calculation of E0INT25/E0EXT25 ##
##     new sf_calib function      ##
####
calib_spectro <- sf_calib(calEint = d_all$voltINT, calEext = d_all$voltEXT, calpH = d_all$ph_s_sf_temp_insi, calT = d_all$temp_sf, calSal = d_all$sal_insitu_ctd_interp)
# put number of deployment 
calib_spectro$deployment_sf <- d_all$deployment_sf

# make the mean of E0 by period of deployement
calib_spectro <- group_by(calib_spectro, deployment_sf) %>%
  summarise(E0int25 = mean(E0int25, na.rm = TRUE),
            E0ext25 = mean(E0ext25, na.rm = TRUE) 
            )
names(calib_spectro) <- c("deployment_sf", "E0int25_sf", "E0ext25_sf")

# add E0 mean to seaFET data per deployment
d_all <-left_join(d_all, calib_spectro,by="deployment_sf" )

#### plot
kk <- tidyr::pivot_longer(calib_spectro, cols = starts_with("E0"), names_to = "variable", values_to = "value")
p_calib <- ggplot(kk) + geom_point(aes(x=deployment_sf, y=value, color=variable), size=2) +
  facet_grid(variable ~ ., scales="free_y") + theme_bw() +
  scale_colour_discrete(guide="none")
print(p_calib)
```

### **seaFET pHT calibrated**
pH is measured continuously by an autonomous seaFET pH sensor. Final calibrated seaFET pHT data (green dots) are shown here after calibration with spectro measurements (red dots) according to each period of deployement, defined between 2 spectro measurements (shading grey).
Calibration was done following the methods and code described in Bresnahan *et al.* (2014) using seaFET voltage and spectro measurements.
During the fifth period, seaFET cannot be calibrated because seaFET was not working during the time of reference (pH spectro) sampling (2018-04-16). To solve this calibration, we decided to fit the spectro measurement with a pH mean of 4 previous days.

@@

- pH INT - pH EXT was plotted here.
